---
title: "Project Proposal"
author: "Las Tortugueras"
date: "02/11/2023"
output: github_document
---

## Load Packages
```{r load_packages, message = FALSE}
install.packages("tidyverse")
install.packages("skimr")

library(tidyverse)
library(skimr)
```

## 1. Introduction

The aim of this project is to analyze data collected on sea turtles in Kino Bay, Sonora, Mexico and Buck Island, Saint Croix, USVI. This project will focus on data collected by Grupo Tortuguero de Bahia de Kino (GTBK), a community organization that monitors local sea turtle populations through boat surveys and mark-recapture. We will work alongside the community organization to answer questions that have come up during their research. Our aim is to address questions that have been raised by GTBK and offer them an analysis of their data. We will examine registration and recapture data including season of registration and frequency of recapture. We will also incorporate data from the Buck Island Sea Turtle Research Program (BISTRP) that monitors nesting sea turtles on an island in the Caribbean. The datasets have individual sea turtles as the case subject and both research projects have collected morphometric data on reproductive female Green sea turtles. Using the overlapping size variables, we will be able to compare the reproductive female Green turtles in Kino Bay, Mexico to those on Buck Island, USVI.

We will work with three datasets: two from Kino Bay (one from 2014-2018 and one from 2019-present) and one from Buck Island.

The variables are as follows:

"Kino Bay 2014-2018" - `ID#`, `Nombre de la tortuga`, `No. Monitoreo`,`Estación`, `Fecha Area de Monitoreo`, `Posicion (Lati. Long.)`, `Hora de capture`, `Especie`, `Peso(Kg)`, `Sexo`, `L.R.C.(cm)`, `L.C.C.(cm)`, `A.R.C.(cm)`, `A.C.C.(cm)`, `P.C (cm)`, `L.P (cm)`, `LTC (cm)`, `¿Posee marca? (Recaptura)`, `Marca Derecha`, `Marca Izquierda`, `Observaciones y comentarios`

"Kino Bay 2019-current" - `Codigo_Comunidad Responsable`, `Red numero`, `Fecha_inicio`, `Hora_inicio`, `Hora_fin`, `Total_horas`, `Tipo_monitoreo`, `Metodologia`, `Longitud red / lanceos`, `Lugar_captura	(Latitud, Longitud)`, `Unidad (UTM o grados)`, `Fecha_captura`, `Hora_captura`, `Nombre tortuga`, `Especie`, `LRC`, `ARC`, `LCC`, `ACC`, `Profundidad cuerpo`, `Longitud plastron`, `Longitud total cola`, `Peso_kg`, `Sexo`, `Material_marcas`, `Marca nueva DX`, `Marca nueva SX`, `Marca previa DX`, `Marca previa SX`, `PIT tag (si, no)`, `PIT nuevo`, `PIT previo`, `Muestra sangre (si o no)`, `Muestra piel(si o no)`, `Muestra cont estom  (si o no)`, `Otra muestra  (si o no)`, `Especificar`, `Lesiones / Balanos`, `Foto (si o no)`, `Comentarios`, `ID #`

"Buck Island 2022" - `activity number`, `CDate`, `PrimaryTag`, `sp`, `ApplyDate`, `Remi?`, `cclnucn`, `ccw`, `cclnuct`, `cclnott`


## 2. Data

```{r load-data, message = FALSE}
library(readxl)
GTBK_2018_2023 <- read_excel("../data/GTBK_2018_2023.xlsx") 

library(readxl)
GTBK_2014_2018 <- read_excel("../data/GTBK_2014_2018.xls")

library(readr)
buis_2022 <- read_csv("../data/buis_greenturtles_2022.csv")
```

#### Data Skim 

```{r skim_data}
skim(GTBK_2018_2023)
skim(GTBK_2014_2018)
skim(buis_2022)
```

## 3. Data Analysis Plan

We are working alongside the community organization Grupo Tortuguero de Bahia de Kino (GTBK). They are a community-based turtle monitoring group that collects data from local estuaries and less frequently from local islands. They are monitoring the health and size of sea turtle populations, and observing potential recapture. GTBK would like to visualize the number of sea turtles that have been registered, number registered per effort hour across all years, frequency of recapture, the average time between recaptures, and the seasonality of recaptures. The data we are using has been collected by GTBK from 2014-present in Laguna la Cruz estuary in Kino Bay, Sonora, Mexico and Isla San Pedro Martir oceanic offshore island in the Gulf of California.

We would also like to incorporate some of the data that Georgia helped collect last summer and fall on Buck Island, Saint Croix, USVI. This data was collected as part of a long term NPS sea turtle monitoring project called the Buck Island Sea Turtle Research Program (BISTRP). BISTRP has been monitoring the nesting sea turtles (Greens, Hawksbills) on Buck Island for 35 years, conducting a mark-recapture study, collecting biological data and recording the locations of nesting activities. We are interested in making some sort of comparison between the reproductive female Green sea turtles (this is one parameter that both of the datasets have in common) sampled on Buck Island, USVI and those sampled in Kino Bay, Sonora, Mexico. Because both research projects collected Curved Carapace Length (CCL) and Curved Carapace Width (CCW) as size measurements of reproductive female Greens, we would like to ask how the size of reproductive female GR sea turtles varies geographically. 

For the GTBK data the subject cases are each an individual sea turtle. The variables we will analyze include date, season, location, tag numbers, species, and sex. 

For the BISTRP data we want to use curved carapace length, curved carapace width, location, species and primary tag number.

#### Preliminary Analysis
```{r clean_species_variable}
GTBK_2018_2023 <- GTBK_2018_2023 %>%
  mutate(especie_nombre_latino = 
           case_when(Especie >= "Prieta/Verde" ~ "Chelonia mydas",
                     Especie >= "Carey" ~ "Eretmochelys imbricata", 
                     Especie >= "Amarilla" ~ "Caretta caretta", 
                     Especie >= "Golfina" ~ "Lepidochelys olivacea"))

```

```{r buis-summary-stats}
summary(buis_2022)
```

```{r GTBK-2014-2018-summary-stats}
summary(GTBK_2014_2018)
```

```{r GTBK-2018-2023-summary-stats}
summary(GTBK_2018_2023)
```

```{r species-breakdown}
GTBK_2014_2018 %>%
  count(Especie, sort = TRUE)

GTBK_2018_2023 %>%
  count(especie_nombre_latino, sort = TRUE)

buis_2022 %>%
  count(sp, sort = TRUE)
```

## Visualizations 

Brainstorming: Some visualizations that could be helpful for our data analysis include line graphs or a time series that maps the number of recaptures by GTBK over time (maybe using some sort of time-based animation). One visualization that we could create would display the number of recaptures by GTBK per species faceted by seasonality. For the visualization comparing the size of reproductive female GR turtles, we could use a pair of violin plots or box plots to display the distribution of individual sizes in each geographic region.

- A bar graph or line graph (or combo) displaying turtles per effort hour over year
- A bar graph (sideways) displaying number of turtles that have been recaptured 2, 3, 4, 5, etc. times
- Spatial color-coded points for turtles that have been recaptured more than once (site fidelity visualization) in locations
- Number of turtles recaptured every year, number recaptured every two years, number recaptured every three years, etc.
- A horizontal bar graph displaying Average number of turtle recaptures over month 
- Bar graphs of number of recaptures over month faceted by individual turtles (2 sets of this for San Pedro Martí and the estuary)
